---
title: "Basic Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```
# Basic Example ## Installation This package is still under development, so we'll install it locally from source. I recommend the R `pak` package. It's becoming the default, is very nice, and handles dependencies well. Then, you can install the `OncoProfilingTools` package from the local source directory like this:
```{r eval=FALSE}
if (!requireNamespace("pak", quietly = TRUE)) {
  install.packages("pak")
}

pak::pkg_install(".", ask = FALSE)

library(OncoProfilingTools)
```

You're now ready to create your first experiment! If you're familiar with the common Seurat library, an `OncoExperiment` is modeled after a `SeuratObject` and behaves similarly. A new blank experiment requires no set arguments!
```{r eval=FALSE}
exp <- OncoExperiment(project_name = "My experiment!")
exp
```

If you have the required assays downloaded, you can attach them now or download them.
```{r eval=FALSE}
# OncoProfilingTools::download_assays(exp, assay_type = "Expression")

OncoProfilingTools::load_assays(exp, "Expression")
```
```{r eval=FALSE}
# OncoProfilingTools::download_assays(exp, assay_type = "Expression")
styler::style_pkg()
roxygen2::roxygenize(clean = TRUE)
pak::pkg_install(".", ask = FALSE)

library(MultiAssayExperiment)
library(SummarizedExperiment)
library(S4Vectors)

library(OncoProfilingTools)

exp <- OncoExperiment(project_name = "Khader Project")
exp <- load_assays(exp, AssayTypes$Expression)
exp_coad <- subset(exp, DepmapModelType == "COAD")
# Future steps:
plots()
analyses()
export()

exp <- OncoExperiment(project_name = "My experiment!")
exp <- load_assays(exp, AssayTypes$Expression)
exp
plot_donut(exp, assay = AssayTypes$Expression, variable = ModelMetadataFields$PatientRace)
```
## Navigating `colData` and assay metadata `OncoExperiment` is a `MultiAssayExperiment`, so there are two places to look when you want model metadata: - `colData(exp)` for the metadata attached to the `OncoExperiment` itself - `colData(experiments(exp)[["Expression"]])` for the metadata attached to the loaded Expression assay If a field like `DepmapModelType`, `ModelID`, or `PatientRace` shows up as `NULL`, first inspect the available column names: ```r colnames(colData(exp)) ``` Once you confirm the column exists, you can filter directly: ```r colData(exp)$DepmapModelType == "COAD" ``` You can also work with the assay-level metadata and its aligned columns: ```r expression_assay <- experiments(exp)[["Expression"]] colnames(colData(expression_assay)) rownames(colData(expression_assay)) == colnames(expression_assay) ``` That alignment is what keeps the loaded model metadata attached to the expression columns, so it is safe to use for filtering and subsetting.
<!-- R Notebook saved outputs: do not edit
RNOTEBOOK_RMD_STATE_BEGIN
eyJ2ZXJzaW9uIjoyLCJjb2RlQ2VsbFN0YXRlcyI6W3siZmluZ2VycHJpbnQiOiJ7XCJsYW5ndWFnZVwiOlwiclwiLFwiY29kZVwi
Olwia25pdHI6Om9wdHNfY2h1bmskc2V0KGV2YWwgPSBGQUxTRSlcIixcIm9wdGlvbnNcIjp7XCJpbmNsdWRlXCI6ZmFsc2UsXCJs
YWJlbFwiOlwic2V0dXBcIn19IiwicmVzdWx0IjpudWxsfSx7ImZpbmdlcnByaW50Ijoie1wibGFuZ3VhZ2VcIjpcInJcIixcImNv
ZGVcIjpcImlmICghcmVxdWlyZU5hbWVzcGFjZShcXFwicGFrXFxcIiwgcXVpZXRseSA9IFRSVUUpKSB7XFxuICBpbnN0YWxsLnBh
Y2thZ2VzKFxcXCJwYWtcXFwiKVxcbn1cXG5cXG5wYWs6OnBrZ19pbnN0YWxsKFxcXCIuXFxcIiwgYXNrID0gRkFMU0UpXFxuXFxu
bGlicmFyeShPbmNvUHJvZmlsaW5nVG9vbHMpXCIsXCJvcHRpb25zXCI6e1wiZXZhbFwiOmZhbHNlfX0iLCJyZXN1bHQiOm51bGx9
LHsiZmluZ2VycHJpbnQiOiJ7XCJsYW5ndWFnZVwiOlwiclwiLFwiY29kZVwiOlwiZXhwIDwtIE9uY29FeHBlcmltZW50KHByb2pl
Y3RfbmFtZSA9IFxcXCJNeSBleHBlcmltZW50IVxcXCIpXFxuZXhwXCIsXCJvcHRpb25zXCI6e1wiZXZhbFwiOmZhbHNlfX0iLCJy
ZXN1bHQiOm51bGx9LHsiZmluZ2VycHJpbnQiOiJ7XCJsYW5ndWFnZVwiOlwiclwiLFwiY29kZVwiOlwiIyBPbmNvUHJvZmlsaW5n
VG9vbHM6OmRvd25sb2FkX2Fzc2F5cyhleHAsIGFzc2F5X3R5cGUgPSBcXFwiRXhwcmVzc2lvblxcXCIpXFxuXFxuT25jb1Byb2Zp
bGluZ1Rvb2xzOjpsb2FkX2Fzc2F5cyhleHAsIFxcXCJFeHByZXNzaW9uXFxcIilcIixcIm9wdGlvbnNcIjp7XCJldmFsXCI6ZmFs
c2V9fSIsInJlc3VsdCI6bnVsbH0seyJmaW5nZXJwcmludCI6IntcImxhbmd1YWdlXCI6XCJyXCIsXCJjb2RlXCI6XCIjIE9uY29Q
cm9maWxpbmdUb29sczo6ZG93bmxvYWRfYXNzYXlzKGV4cCwgYXNzYXlfdHlwZSA9IFxcXCJFeHByZXNzaW9uXFxcIilcXG5cXG5w
YWs6OnBrZ19pbnN0YWxsKFxcXCIuXFxcIiwgYXNrID0gRkFMU0UpXFxuXFxubGlicmFyeShNdWx0aUFzc2F5RXhwZXJpbWVudClc
XG5saWJyYXJ5KFN1bW1hcml6ZWRFeHBlcmltZW50KVxcbmxpYnJhcnkoUzRWZWN0b3JzKVxcblxcbmxpYnJhcnkoT25jb1Byb2Zp
bGluZ1Rvb2xzKVxcblxcbmV4cCA8LSBPbmNvRXhwZXJpbWVudChwcm9qZWN0X25hbWUgPSBcXFwiS2hhZGVyIFByb2plY3RcXFwi
KVxcbmV4cCA8LSBsb2FkX2Fzc2F5cyhleHAsIEFzc2F5VHlwZXMkRXhwcmVzc2lvbilcXG5leHBfY29hZCA8LSBzdWJzZXQoZXhw
LCBEZXBtYXBNb2RlbFR5cGUgPT0gXFxcIkNPQURcXFwiKVxcbiMgRnV0dXJlIHN0ZXBzOlxcbnBsb3RzKClcXG5hbmFseXNlcygp
XFxuZXhwb3J0KClcXG5cXG5cXG4jIGdldCBhbGwgUFJJU00gZHJ1ZyByZXNwb25zZSBkYXRhIHdoZXJlIHRoZSBPbmNvVHJlZUxp
bmVhZ2UgaXMgYm93ZWxcXG5cXG5wcmlzbSA8LSBleHBlcmltZW50cyhleHApJFxcXCJQUklTTVxcXCJcXG5wcm90ZWluIDwtIGV4
cGVyaW1lbnRzKGV4cClbW1xcXCJQcm90ZWluIEV4cHJlc3Npb25cXFwiXV1cXG5cXG5wcm90ZWluX2NvbGRhdGEgPC0gYXMuZGF0
YS5mcmFtZShjb2xEYXRhKHByb3RlaW4pKVxcblxcbmJvd2VsX21vZGVsX2lkcyA8LSByb3duYW1lcyhwcm90ZWluX2NvbGRhdGEp
W1xcbiAgcHJvdGVpbl9jb2xkYXRhJE9uY290cmVlTGluZWFnZSA9PSBcXFwiQm93ZWxcXFwiXFxuXVxcblxcbmJvd2VsX21vZGVs
X2lkcyA8LSBpbnRlcnNlY3QoXFxuICBib3dlbF9tb2RlbF9pZHMsXFxuICBjb2xuYW1lcyhwcmlzbSlcXG4pXFxuXFxucHJpc21f
Ym93ZWwgPC0gcHJpc21bLCBib3dlbF9tb2RlbF9pZHNdXFxucHJpc21fYm93ZWxfbWF0cml4IDwtIGFzc2F5KHByaXNtX2Jvd2Vs
LCBcXFwicmVzcG9uc2VcXFwiKVxcbmRpbShwcmlzbV9ib3dlbF9tYXRyaXgpXFxuXFxucHJpc21fYm93ZWwgIyBub3RpY2Ugd2Ug
aGF2ZSBvbmx5IHRoZSA0NSBib3dlbGFuYWx5c2VzKClcIixcIm9wdGlvbnNcIjp7XCJldmFsXCI6ZmFsc2V9fSIsInJlc3VsdCI6
eyJ0eXBlIjoicmVzdWx0IiwiY2h1bmtfaWQiOiJjaHVuay04Iiwic291cmNlX2NvZGUiOiIjIE9uY29Qcm9maWxpbmdUb29sczo6
ZG93bmxvYWRfYXNzYXlzKGV4cCwgYXNzYXlfdHlwZSA9IFwiRXhwcmVzc2lvblwiKVxuXG5wYWs6OnBrZ19pbnN0YWxsKFwiLlwi
LCBhc2sgPSBGQUxTRSlcblxubGlicmFyeShNdWx0aUFzc2F5RXhwZXJpbWVudClcbmxpYnJhcnkoU3VtbWFyaXplZEV4cGVyaW1l
bnQpXG5saWJyYXJ5KFM0VmVjdG9ycylcbmxpYnJhcnkoT25jb1Byb2ZpbGluZ1Rvb2xzKVxuXG5cbmV4cCA8LSBPbmNvRXhwZXJp
bWVudChwcm9qZWN0X25hbWUgPSBcIktoYWRlciBQcm9qZWN0XCIpXG5leHAgPC0gbG9hZF9hc3NheXMoZXhwLCBBc3NheVR5cGVz
JEV4cHJlc3Npb24pXG5leHBfY29hZCA8LSBzdWJzZXQoZXhwLCBEZXBtYXBNb2RlbFR5cGUgPT0gXCJDT0FEXCIpXG4jIEZ1dHVy
ZSBzdGVwczpcbnBsb3RzKClcbiMgZ2V0IGFsbCBQUklTTSBkcnVnIHJlc3BvbnNlIGRhdGEgd2hlcmUgdGhlIE9uY29UcmVlTGlu
ZWFnZSBpcyBib3dlbFxuXG5wcmlzbSA8LSBleHBlcmltZW50cyhleHApJFwiUFJJU01cIlxucHJvdGVpbiA8LSBleHBlcmltZW50
cyhleHApW1tcIlByb3RlaW4gRXhwcmVzc2lvblwiXV1cblxucHJvdGVpbl9jb2xkYXRhIDwtIGFzLmRhdGEuZnJhbWUoY29sRGF0
YShwcm90ZWluKSlcblxuYm93ZWxfbW9kZWxfaWRzIDwtIHJvd25hbWVzKHByb3RlaW5fY29sZGF0YSlbXG4gIHByb3RlaW5fY29s
ZGF0YSRPbmNvdHJlZUxpbmVhZ2UgPT0gXCJCb3dlbFwiXG5dXG5cbmJvd2VsX21vZGVsX2lkcyA8LSBpbnRlcnNlY3QoXG4gIGJv
d2VsX21vZGVsX2lkcyxcbiAgY29sbmFtZXMocHJpc20pXG4pXG5cbnByaXNtX2Jvd2VsIDwtIHByaXNtWywgYm93ZWxfbW9kZWxf
aWRzXVxucHJpc21fYm93ZWxfbWF0cml4IDwtIGFzc2F5KHByaXNtX2Jvd2VsLCBcInJlc3BvbnNlXCIpXG5kaW0ocHJpc21fYm93
ZWxfbWF0cml4KVxuXG5wcmlzbV9ib3dlbCAjIG5vdGljZSB3ZSBoYXZlIG9ubHkgdGhlIDQ1IGJvd2VsIiwiY29uc29sZV9zZWdt
ZW50cyI6W3siY29kZSI6IiMgT25jb1Byb2ZpbGluZ1Rvb2xzOjpkb3dubG9hZF9hc3NheXMoZXhwLCBhc3NheV90eXBlID0gXCJF
eHByZXNzaW9uXCIpXG5cbnBhazo6cGtnX2luc3RhbGwoXCIuXCIsIGFzayA9IEZBTFNFKVxuXG5saWJyYXJ5KE11bHRpQXNzYXlF
eHBlcmltZW50KVxubGlicmFyeShTdW1tYXJpemVkRXhwZXJpbWVudClcbmxpYnJhcnkoUzRWZWN0b3JzKVxubGlicmFyeShPbmNv
UHJvZmlsaW5nVG9vbHMpXG5cblxuZXhwIDwtIE9uY29FeHBlcmltZW50KHByb2plY3RfbmFtZSA9IFwiS2hhZGVyIFByb2plY3Rc
IilcbmV4cCA8LSBsb2FkX2Fzc2F5cyhleHAsIEFzc2F5VHlwZXMkRXhwcmVzc2lvbilcbmV4cF9jb2FkIDwtIHN1YnNldChleHAs
IERlcG1hcE1vZGVsVHlwZSA9PSBcIkNPQURcIilcbiMgRnV0dXJlIHN0ZXBzOlxucGxvdHMoKVxuIyBnZXQgYWxsIFBSSVNNIGRy
dWcgcmVzcG9uc2UgZGF0YSB3aGVyZSB0aGUgT25jb1RyZWVMaW5lYWdlIGlzIGJvd2VsXG5cbnByaXNtIDwtIGV4cGVyaW1lbnRz
KGV4cCkkXCJQUklTTVwiXG5wcm90ZWluIDwtIGV4cGVyaW1lbnRzKGV4cClbW1wiUHJvdGVpbiBFeHByZXNzaW9uXCJdXVxuXG5w
cm90ZWluX2NvbGRhdGEgPC0gYXMuZGF0YS5mcmFtZShjb2xEYXRhKHByb3RlaW4pKVxuXG5ib3dlbF9tb2RlbF9pZHMgPC0gcm93
bmFtZXMocHJvdGVpbl9jb2xkYXRhKVtcbiAgcHJvdGVpbl9jb2xkYXRhJE9uY290cmVlTGluZWFnZSA9PSBcIkJvd2VsXCJcbl1c
blxuYm93ZWxfbW9kZWxfaWRzIDwtIGludGVyc2VjdChcbiAgYm93ZWxfbW9kZWxfaWRzLFxuICBjb2xuYW1lcyhwcmlzbSlcbilc
blxucHJpc21fYm93ZWwgPC0gcHJpc21bLCBib3dlbF9tb2RlbF9pZHNdXG5wcmlzbV9ib3dlbF9tYXRyaXggPC0gYXNzYXkocHJp
c21fYm93ZWwsIFwicmVzcG9uc2VcIilcbmRpbShwcmlzbV9ib3dlbF9tYXRyaXgpXG5cbnByaXNtX2Jvd2VsICMgbm90aWNlIHdl
IGhhdmUgb25seSB0aGUgNDUgYm93ZWwiLCJvdXRwdXQiOiIifV0sImNvbnNvbGUiOiIiLCJzdGRvdXQiOiIiLCJzdGRlcnIiOiIi
LCJwbG90cyI6W10sInBsb3RzX2h0bWwiOltdLCJkYXRhZnJhbWVzIjpbXSwiZXJyb3IiOiJSIHByb2Nlc3MgZXhpdGVkIChjb2Rl
PTEpIn19XSwiY29kZUNlbGxSZXN1bHRzIjpbXX0=
RNOTEBOOK_RMD_STATE_END
-->
