---
title: "Basic Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```
# Basic Example ## Installation This package is still under development, so we'll install it locally from source. I recommend the R `pak` package. It's becoming the default, is very nice, and handles dependencies well. Then, you can install the `OncoProfilingTools` package from the local source directory like this:
```{r eval=FALSE}
if (!requireNamespace("pak", quietly = TRUE)) {
  install.packages("pak")
}

pak::pkg_install(".", ask = FALSE)

library(OncoProfilingTools)
```

You're now ready to create your first experiment! If you're familiar with the common Seurat library, an `OncoExperiment` is modeled after a `SeuratObject` and behaves similarly. A new blank experiment requires no set arguments!
```{r eval=FALSE}
exp <- OncoExperiment(project_name = "My experiment!")
exp
```

If you have the required assays downloaded, you can attach them now or download them.
```{r eval=FALSE}
# OncoProfilingTools::download_assays(exp, assay_type = "Expression")

OncoProfilingTools::load_assays(exp, "Expression")
```
```{r eval=FALSE}
# OncoProfilingTools::download_assays(exp, assay_type = "Expression")

pak::pkg_install(".", ask = FALSE)

library(MultiAssayExperiment)
library(SummarizedExperiment)
library(S4Vectors)
library(OncoProfilingTools)


exp <- OncoExperiment(project_name = "My experiment!")
exp <- load_assays(exp, AssayTypes$Expression)
exp <- load_assays(exp, AssayTypes$ProteinExpression)
exp <- load_assays(exp, AssayTypes$PRISM)
exp

# get all PRISM drug response data where the OncoTreeLineage is bowel

prism <- experiments(exp)$"PRISM"
protein <- experiments(exp)[["Protein Expression"]]

protein_coldata <- as.data.frame(colData(protein))

bowel_model_ids <- rownames(protein_coldata)[
  protein_coldata$OncotreeLineage == "Bowel"
]

bowel_model_ids <- intersect(
  bowel_model_ids,
  colnames(prism)
)

prism_bowel <- prism[, bowel_model_ids]
prism_bowel_matrix <- assay(prism_bowel, "response")
dim(prism_bowel_matrix)

prism_bowel # notice we have only the 45 bowel
```
